The regularity of ODEs and thimble integrals \\ with respect to Borel summation \begin{draft}\textbf{[DRAFT]}\end{draft}

### 2024-07-01 exp-resurgence.tex
Try to patch thimble integral resurgence remark

\[ \singexpalg{\tau_\alpha}(\Omega_\alpha) \to e^{-\alpha z} \dualsingexp{-\tau_\alpha-1}(\widehat{\Omega}_\alpha^{\textcolor{LimeGreen}{\bullet}}). \]

@@

\begin{draft}Recall the setting from Section~\ref{borel-reg:explanatory-power}: let $X$ be a $1$-dimensional complex manifold equipped with a volume form $\nu \in \Omega^{1,0}(X)$ \begin{todo}[this originally said $\Gamma(X,\Omega^1)$, but shouldn't it be either $\Omega^{1,0}(X)$ or $\Gamma\big(X, \Lambda^{1,0} (X)\big)$?]\end{todo} and a holomorphic function $f\colon X\to\C$ with non--degenerate critical points. Let $I_a$ be the thimble integral 

@@

where $\mathcal{C}_a^\theta$ is the thimble through the critical point $a$. Let $\alpha$ be the associated critical value, $f(a)$. In this section, we will prove that thimble integrals of the form \eqref{exp-int} are Borel regular. The first step will be to rewrite $I_a$ as the Laplace transform of a function $\iota_a$ which is explicitly given by the well-known formula that we will call the \textit{thimble projection formula}.\end{draft}

@@

where $\mathcal{C}_a^\theta(\zeta)$ is the part of $\mathcal{C}_a^\theta$ that goes through $f^{-1}([\alpha,\zeta e^{i\theta}])$. Notice that $\mathcal{C}_a^\theta(\zeta)$ starts and ends in $f^{-1}(\zeta)$. \begin{todo}[The thimble $\mathcal{C}_a^\theta$ needs an orientation, but the orientation is arbitrary.]\end{todo}

@@

    & = \fracderiv{-1}{\zeta}{\alpha} \partial_\zeta \left( \int_{\mathcal{C}_a^\theta(\zeta)}\nu \right) \\

@@

@@

We will calculate the jump of $\int_{\mathcal{C}_a^\theta(\zeta)}\nu$ across the cut starting at $\zeta = \beta$. Write $\beta$ as $\alpha + re^{i\theta_\beta}$, and let $\beta^+$ and $\beta^-$ be the points we get by infinitesimally increasing or decreasing $\theta_\beta$. The jump is then given by \[\int_{\mathcal{C}_a^\theta(\beta^+)} \nu - \int_{\mathcal{C}_a^\theta(\beta^-)} \nu\] 
and it turns out to be proportional to $\series{\phi}_\beta$, namely to the germ of analytic functions at $\zeta_\alpha=\beta$. %We will see an example of this behaviour in Appendix~\ref{apx:eye-res-airy}.

\begin{revised}Initially, the function $\int_{\mathcal{C}^\theta_a(\zeta)} \nu$ is only defined for $\zeta$ on the ray $[\alpha, \infty)$, but it can be analytically continued to points off the ray by homotopy of the path $\mathcal{C}^\theta_a(\zeta)$. Since $\nu$ is holomorphic, the only way for $\int_{\mathcal{C}^\theta_a(\zeta)} \nu$ to be singular is for something to go wrong with the hotmotopy of $\mathcal{C}^\theta_a(\zeta)$. At a critical point of $f$, the ends of the homotoped path $\mathcal{C}^\theta_a(\zeta)$ come together, telling us that \begin{todo}[\textbf{A} \textbf{V} \ldots]\end{todo}\end{revised}
\end{remark}
\begin{todo}
\begin{corollary}\label{simple-res-thimble}
If $f$ and $\nu$ are polynomials, the function ${\phi}=\fracderiv{-1/2}{\zeta}{\alpha}\iota$ defined in Remark~\ref{rmk:1/2-deriv} has simple resurgent singularities at $\zeta=\beta$, for all $\beta$ roots of $P$ such that $\beta\neq\alpha$.  
\end{corollary}
\begin{proof}
Recall ${\phi}_\alpha$ is the sum of the Borel transform of $z^{1/2}\series{I}_j$, namely 
\[\hat{\phi}_j(\zeta)=\fracderiv{1/2}{\zeta}{\alpha_j}\hat{\iota}_j=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\hat{\iota}_j(s)\, ds\]
Then by Pham's result (see \eqref{rmk:Pham formula}) we know 
\[\hat{\iota}_j(\zeta_j)=\frac{(-\zeta_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-\zeta_j)^{-k} + \text{ hol. funct. }\]
hence 
\[\hat{\iota}_j(\zeta)=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\frac{(-s+\alpha_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-s+\alpha_j)^{-k} \, ds + \text{ hol. funct. }\]
\end{proof}
\end{todo}
\begin{remark}\label{rmk:Pham formula}
When $f$ and $\nu$ are polynomials and $X$ is $N$-dimensional, a result of Pham \cite[Equation 2.4, II partie]{pham} gives us an explicit expression for the singularity of the function $\iota_a$ from Lemma~\ref{lem:thimble_proj_formula-proof}. For some coefficients $c_k$ which depend on $f$, $\nu$, and the thimble $\mathcal{C}_a^\theta$, we have
\begin{equation}\label{eqn:Pham}
\iota_a \in (-1)^{\frac{N(N-1)}{2}}  \sum_{k=0}^\infty c_k\, \delta_{\zeta, \alpha}^{(-N/2 - k)} + \mathcal{O}_{\zeta = \alpha},
\end{equation}
where the singularities
\begin{equation*}
\delta_{\zeta,\alpha}^{(-\ell)}:=\begin{cases}
\displaystyle\frac{\zeta_\alpha^{\ell-1}}{-2\pi i(\ell-1)!}\log(\zeta_\alpha)+ \mathcal{O}_{\zeta = \alpha} & \text{ if } \ell\in\mathbb{N}^*\\
& \\
\displaystyle\frac{(-\zeta_\alpha)^{\ell-1}}{2\pi i\, \Gamma(-\ell)}+ \mathcal{O}_{\zeta = \alpha} & \text{ if } \ell\notin \mathbb{N}^* \,.
\end{cases}
\end{equation*}
can be seen as fractional derivatives of the Dirac delta distribution. In particular, if $N=1$, the inverse Laplace transform of the function $I_a$ from Equation~\eqref{exp-int} has a singularity of the form $\zeta_\alpha^{1/2}$, as we compute explicitly in the proof of Theorem~\ref{thm:maxim-proof}. %as computed in Remark~\ref{rmk:1/2-deriv} the inverse Laplace tranform of $z^{-1/2} I$ is the $1/2$-derivative of ${\iota}$, hence it has \textcolor{orange}{simple singularities (see Corollary~\ref{simple-res-thimble})}. \textcolor{magenta}{We expect this result to hold beyond the polynomial case.} We also expect that equation \eqref{eqn:formula-proof} can be generalized to the $N$-dimensional case replacing the $1/2$-derivative with the $N/2$ one. 
\end{remark}
\section{Examples}\label{sec:examples}
%
\subsection{Airy}\label{sec:airy}
%
The Airy equation is
\begin{equation}\label{eqn:airy}
\left[\big(\tfrac{\partial}{\partial y}\big)^2 - y\right] \Phi = 0.
\end{equation}
One solution is given by the Airy function,
\[ \Ai(y) = \frac{1}{2\pi i} \int_\gamma \exp\left(\tfrac{1}{3}t^3 - yt\right)\,dt, \]
where $\gamma$ is a path that comes from $\infty$ at $-60^\circ$ and returns to $\infty$ at $60^\circ$.

>>>>

\[ \singexpalg{\tau_\alpha}(\Omega_\alpha) \to e^{-\alpha z} \dualsingexp{-\tau_\alpha-1}(\widehat{\Omega}_\alpha^\bullet). \]

@@

\begin{revised}Let us recall the setting from Section~\ref{borel-reg:explanatory-power}. Let $X$ be a $1$-dimensional complex manifold equipped with a volume form $\nu \in \Omega^{1,0}(X)$ and a holomorphic function $f\colon X\to\C$ with non--degenerate critical points. Let $I_a$ be the thimble integral 

@@

where $\mathcal{C}_a^\theta$ is the thimble through the critical point $a$. Let $\alpha$ be the associated critical value, $f(a)$. In this section, we will prove that thimble integrals of the form \eqref{exp-int} are Borel regular. The first step is to rewrite $I_a$ as the Laplace transform of a function $\iota_a$, which is given explicitly by the well-known formula that we will call the \textit{thimble projection formula}.\end{revised}

@@

where $\mathcal{C}_a^\theta(\zeta)$ is the part of $\mathcal{C}_a^\theta$ that goes through $f^{-1}([\alpha,\zeta e^{i\theta}])$. Notice that $\mathcal{C}_a^\theta(\zeta)$ starts and ends in $f^{-1}(\zeta)$.

@@

& = \fracderiv{-1}{\zeta}{\alpha} \frac{\partial}{\partial \zeta} \left( \int_{\mathcal{C}_a^\theta(\zeta)}\nu \right) \\

@@

\begin{revised}We can use the thimble projection formula to show that each function $\iota_\alpha$ is resurgent, in the sense of \'{E}calle~\cite[Section~1]{EcalleI}. Let $\mathcal{J}_\alpha(\beta)$ be the straight path from $\zeta = \alpha$ to $\zeta = \beta$ in $\C$. As long as this path avoids the critical values of $f$, it lifts uniquely along $f$ to a path $\mathcal{C}_a^\theta(\beta)$ in $X$. This lets us analytically continue $\int_{\mathcal{C}_a^\theta(\zeta)} \nu$ to a star-shaped domain $\Omega_\alpha \subset \C$. Intuitively, $\Omega_\alpha$ is constructed by drawing rays from $\zeta = \alpha$ through all of the other critical values, and then cutting $\C$ along the parts of those rays that go from the other critical values to $\zeta = \infty$, as shown in Figure~\ref{Fig:slit domain}.\end{revised}

@@

\begin{revised}Initially, the function $\int_{\mathcal{C}_a(\zeta)} \nu$ is only defined for $\zeta$ on the ray $[\alpha, \infty)$, but it can be analytically continued to points off the ray by homotopy of the path $\mathcal{C}_a(\zeta)$. Since $\nu$ is holomorphic, the only way for $\int_{\mathcal{C}^\theta_a(\zeta)} \nu$ to become singular is for something to go wrong with the hotmotopy of $\mathcal{C}_a(\zeta)$, which can only happen at the critical values of $f$. Therefore, $\mathcal{C}_a(\zeta)$ is endlessly analytically continuable away from the critical values of $f$. Resurgent functions have resurgent derivatives \begin{todo}[check]\end{todo}, so $\iota_\alpha$ is resurgent too.\end{revised}
\begin{todo}[Incorporate this additional work]\end{todo} We will calculate the jump of $\int_{\mathcal{C}_a^\theta(\zeta)}\nu$ across the cut starting at $\zeta = \beta$. Write $\beta$ as $\alpha + re^{i\theta_\beta}$, and let $\beta^+$ and $\beta^-$ be the values we get by infinitesimally increasing or decreasing $\theta_\beta$. The jump is then given by \[\int_{\mathcal{C}_a^\theta(\beta^+)} \nu - \int_{\mathcal{C}_a^\theta(\beta^-)} \nu\] 
and it turns out to be proportional to $\series{\phi}_\beta$, namely to the germ of analytic functions at $\zeta_\alpha=\beta$. %We will see an example of this behaviour in Appendix~\ref{apx:eye-res-airy}.

\begin{todo}[This paragraph is marked as revised]\end{todo} Initially, the function $\int_{\mathcal{C}^\theta_a(\zeta)} \nu$ is only defined for $\zeta$ on the ray $[\alpha, \infty)$, but it can be analytically continued to points off the ray by homotopy of the path $\mathcal{C}^\theta_a(\zeta)$. Since $\nu$ is holomorphic, the only way for $\int_{\mathcal{C}^\theta_a(\zeta)} \nu$ to be singular is for something to go wrong with the hotmotopy of $\mathcal{C}^\theta_a(\zeta)$. At a critical point of $f$, the ends of the homotoped path $\mathcal{C}^\theta_a(\zeta)$ come together, telling us that \begin{todo}[\textbf{A} \textbf{V} \ldots]\end{todo}
\end{old}
\end{remark}
\begin{todo}
\begin{corollary}\label{simple-res-thimble}
If $f$ and $\nu$ are polynomials, the function ${\phi}=\fracderiv{-1/2}{\zeta}{\alpha}\iota$ defined in Remark~\ref{rmk:1/2-deriv} has simple resurgent singularities at $\zeta=\beta$, for all $\beta$ roots of $P$ such that $\beta\neq\alpha$.  
\end{corollary}
\begin{proof}
Recall ${\phi}_\alpha$ is the sum of the Borel transform of $z^{1/2}\series{I}_j$, namely 
\[\hat{\phi}_j(\zeta)=\fracderiv{1/2}{\zeta}{\alpha_j}\hat{\iota}_j=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\hat{\iota}_j(s)\, ds\]
Then by Pham's result (see \eqref{rmk:Pham formula}) we know 
\[\hat{\iota}_j(\zeta_j)=\frac{(-\zeta_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-\zeta_j)^{-k} + \text{ hol. funct. }\]
hence 
\[\hat{\iota}_j(\zeta)=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\frac{(-s+\alpha_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-s+\alpha_j)^{-k} \, ds + \text{ hol. funct. }\]
\end{proof}
\end{todo}
\begin{remark}\label{rmk:Pham formula}
When $f$ and $\nu$ are polynomials and $X$ is $N$-dimensional, a result of Pham \cite[Equation 2.4, II partie]{pham} gives us an explicit expression for the singularity of the function $\iota_a$ from Lemma~\ref{lem:thimble_proj_formula-proof}. For some coefficients $c_k$ which depend on $f$, $\nu$, and the thimble $\mathcal{C}_a^\theta$, we have
\begin{equation}\label{eqn:Pham}
\iota_a \in (-1)^{\frac{N(N-1)}{2}}  \sum_{k=0}^\infty c_k\, \delta_{\zeta, \alpha}^{(-N/2 - k)} + \mathcal{O}_{\zeta = \alpha},
\end{equation}
where the singularities
\begin{equation*}
\delta_{\zeta,\alpha}^{(-\ell)}:=\begin{cases}
\displaystyle\frac{\zeta_\alpha^{\ell-1}}{-2\pi i(\ell-1)!}\log(\zeta_\alpha)+ \mathcal{O}_{\zeta = \alpha} & \text{ if } \ell\in\mathbb{N}^*\\
& \\
\displaystyle\frac{(-\zeta_\alpha)^{\ell-1}}{2\pi i\, \Gamma(-\ell)}+ \mathcal{O}_{\zeta = \alpha} & \text{ if } \ell\notin \mathbb{N}^* \,.
\end{cases}
\end{equation*}
can be seen as fractional derivatives of the Dirac delta distribution. In particular, if $N=1$, the inverse Laplace transform of the function $I_a$ from Equation~\eqref{exp-int} has a singularity of the form $\zeta_\alpha^{1/2}$, as we compute explicitly in the proof of Theorem~\ref{thm:maxim-proof}. %as computed in Remark~\ref{rmk:1/2-deriv} the inverse Laplace tranform of $z^{-1/2} I$ is the $1/2$-derivative of ${\iota}$, hence it has \textcolor{orange}{simple singularities (see Corollary~\ref{simple-res-thimble})}. \textcolor{magenta}{We expect this result to hold beyond the polynomial case.} We also expect that equation \eqref{eqn:formula-proof} can be generalized to the $N$-dimensional case replacing the $1/2$-derivative with the $N/2$ one. 
\end{remark}
\section{Examples}\label{sec:examples}
%
\subsection{Airy}\label{sec:airy}
%
The Airy equation is
\begin{equation}\label{eqn:airy}
\left[\big(\tfrac{\partial}{\partial y}\big)^2 - y\right] \Phi = 0.
\end{equation}
One solution is given by the Airy function,
\[ \Ai(y) = \frac{1}{2\pi i} \int_\gamma \exp\left(\tfrac{1}{3}t^3 - yt\right)\,dt, \]
where $\gamma$ is a path that comes from $\infty$ at $-60^\circ$ and returns to $\infty$ at $60^\circ$.
